Parameters estimation via HOSVD

### 2024-10-23 Code/R/Research/HOSVD parameters estimate.Rmd
Singular values ratio for noise separation

n <- 25
L <- 15
I <- 14; L.tens <- 8

# no rates
# rates.true <- c(0, 0)
# both rates -0.01
# rates.true <- c(-0.01, -0.01)
# both rates -0.02
# rates.true <- c(-0.02, -0.02)
# rates from paper
rates.true <- c(-0.01, -0.02)

@@

R <- 100
l.range <- 2:4

lambda.mean <- list(ssa = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)), 
                    tssa = list(
                      dim1 = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)),
                      dim2 = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)),
                      dim3 = matrix(numeric(length(l.range) * len.out), nrow = length(l.range))))
colnames(lambda.mean$ssa) <- round(sds, 2)
lambda.mean$tssa <- lapply(lambda.mean$tssa, `colnames<-`, value = round(sds, 2))

@@

@@

    lambda.ssa[,i] <- svd(Rssa::hankel(x, L))$d[l.range]

    Z <- hosvd_mod(tens3(x, I, L.tens), status = FALSE)$Z@data
    lambda.tssa.dim1[,i] <- apply(Z, 1, fnorm_complex)[l.range]
    lambda.tssa.dim2[,i] <- apply(Z, 2, fnorm_complex)[l.range]
    lambda.tssa.dim3[,i] <- apply(Z, 3, fnorm_complex)[l.range]

@@

  lambda.mean$ssa[,j] <- rowMeans(lambda.ssa)
  lambda.mean$tssa$dim1[,j] <- rowMeans(lambda.tssa.dim1)
  lambda.mean$tssa$dim2[,j] <- rowMeans(lambda.tssa.dim2)
  lambda.mean$tssa$dim3[,j] <- rowMeans(lambda.tssa.dim3)
}
close(pb)
```

@@

ggsave("./img/Scree_plots_diffrates.pdf",
       marrangeGrob(grobs = scree.plots, ncol = 2, nrow = 3,
                    layout_matrix = layout_matrix,
                    top = quote(
                      paste0("Rates = (", rates.true[1], ", ", rates.true[2],
                                      "), page ", g, " of ", npages))),
       width = 9, height = 9)
```

>>>>

n <- 25
L <- 15
I <- 14
L.tens <- 8

@@

R <- 500
l.range <- 2:4

lambda.mean <- list(
  ssa = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)),
  tssa = list(
    dim1 = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)),
    dim2 = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)),
    dim3 = matrix(numeric(length(l.range) * len.out), nrow = length(l.range))
  )
)

noise.lambda.mean <- list(
  ssa = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)),
  tssa = list(
    dim1 = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)),
    dim2 = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)),
    dim3 = matrix(numeric(length(l.range) * len.out), nrow = length(l.range))
  )
)

colnames(lambda.mean$ssa) <- round(sds, 2)
colnames(noise.lambda.mean$ssa) <- round(sds, 2)
lambda.mean$tssa <- lapply(lambda.mean$tssa, `colnames<-`, value = round(sds, 2))
noise.lambda.mean$tssa <- lapply(noise.lambda.mean$tssa, `colnames<-`, value = round(sds, 2))

@@

  noise.lambda.ssa <- matrix(numeric(length(l.range) * R), nrow = length(l.range))
  noise.lambda.tssa.dim1 <- matrix(numeric(length(l.range) * R), nrow = length(l.range))
  noise.lambda.tssa.dim2 <- matrix(numeric(length(l.range) * R), nrow = length(l.range))
  noise.lambda.tssa.dim3 <- matrix(numeric(length(l.range) * R), nrow = length(l.range))

@@

    lambda.ssa[, i] <- svd(Rssa::hankel(x, L))$d[l.range]

    Z <- hosvd_mod(tens3(x, I, L.tens), status = FALSE)$Z@data
    lambda.tssa.dim1[, i] <- apply(Z, 1, fnorm_complex)[l.range]
    lambda.tssa.dim2[, i] <- apply(Z, 2, fnorm_complex)[l.range]
    lambda.tssa.dim3[, i] <- apply(Z, 3, fnorm_complex)[l.range]

    if(sd > 0) {
      noise.lambda.ssa[, i] <- svd(Rssa::hankel(noise, L))$d[seq_along(l.range)]

      Z.noise <- hosvd_mod(tens3(noise, I, L.tens), status = FALSE)$Z@data
      noise.lambda.tssa.dim1[, i] <- apply(Z.noise, 1, fnorm_complex)[seq_along(l.range)]
      noise.lambda.tssa.dim2[, i] <- apply(Z.noise, 2, fnorm_complex)[seq_along(l.range)]
      noise.lambda.tssa.dim3[, i] <- apply(Z.noise, 3, fnorm_complex)[seq_along(l.range)]
    }

@@

  lambda.mean$ssa[, j] <- rowMeans(lambda.ssa)
  lambda.mean$tssa$dim1[, j] <- rowMeans(lambda.tssa.dim1)
  lambda.mean$tssa$dim2[, j] <- rowMeans(lambda.tssa.dim2)
  lambda.mean$tssa$dim3[, j] <- rowMeans(lambda.tssa.dim3)

  noise.lambda.mean$ssa[, j] <- rowMeans(noise.lambda.ssa)
  noise.lambda.mean$tssa$dim1[, j] <- rowMeans(noise.lambda.tssa.dim1)
  noise.lambda.mean$tssa$dim2[, j] <- rowMeans(noise.lambda.tssa.dim2)
  noise.lambda.mean$tssa$dim3[, j] <- rowMeans(noise.lambda.tssa.dim3)
}
close(pb)
```

@@

ggsave("./img/Scree_plots_norates.pdf",
       marrangeGrob(grobs = scree.plots, ncol = 2, nrow = 3,
                    layout_matrix = layout_matrix,
                    top = quote(
                      paste0("Rates = (", rates.true[1], ", ", rates.true[2],
                                      "), page ", g, " of ", npages))),
       width = 9, height = 9)
```

## Отношения разниц сингулярных значений
```{r, results='hold'}
cat("SSA\n|--Noise ratios:\n")
apply(noise.lambda.mean$ssa[, -1], 2, function(x) Reduce(`/`, -diff(x)))
cat("\n|--Signal+Noise ratios:\n")
apply(lambda.mean$ssa[, -1], 2, function(x) Reduce(`/`, -diff(x)))

cat("\n\nHO-SSA\n|--Noise ratios:\n")
apply(noise.lambda.mean$tssa$dim1[, -1], 2, function(x) Reduce(`/`, -diff(x)))
cat("\n|--Signal+Noise ratios:\n")
apply(lambda.mean$tssa$dim1[, -1], 2, function(x) Reduce(`/`, -diff(x)))
```
